\textbf{Efficient Implementation Strategies for Block Ciphers on ARMv8}\\ {\footnotesize Bachelorarbeit}

### 2023-04-14 Paper/chapters/implementation-strategies.tex
more absorption

>>>>

\subsection{Platform-independent techniques}

The original paper proposes various platform-independent ways to implement
Camellia efficiently. Only some of these apply to the ARMv8 architecture since
features like the inline barrel shifter and bitfield manipulation instructions
generally offer better performance.

\begin{description}
    \item[XOR cancellation property in key schedule:]
        \begin{align*}
            K_A&\leftarrow K_L\oplus K_R \\
            K_A&\leftarrow FE(FE(K_A, \Sigma_0=0xa09e667f3bcc908b), \Sigma_1=0xb67ae8584caa73b2) \\
            K_A&\leftarrow K_A\oplus K_L \\
        \end{align*}

        causes cancellations, allowing us to eliminate some operations by
        replacing it with the following:

        \begin{align*}
            K_{A_L}&\leftarrow F(K_{L_R}\oplus F(K_{L_L}, \Sigma_0)) \\
            K_{A_R}&\leftarrow F(K_{L_L}\oplus \Sigma_1)
        \end{align*}

    \item[Absorption of whitening keys:] Key additions $kw_1,kw_3$ can be
        absorped into other subkeys:

\end{description}

### 2023-04-14 Paper/chapters/introduction.tex
camellia intro FL
more absorption

\subsubsection{Encryption overview}

@@

@@

\paragraph{Feistel round function $FE$}

@@

\paragraph{SP-function $F$}

@@

\paragraph{Substitution function $F$}

\begin{align*}
    S:\F{64} &\rightarrow \F{64} \\
    l_{0(8)||l_{1(8)||l_{2(8)||l_{3(8)||l_{4(8)||l_{5(8)||l_{6(8)||l_{7(8)||&\mapsto P(S(X\oplus k))
\end{align*}

\paragraph{$FL$ layer function $FLL$}

\begin{align*}
    FLL:(\F{64})^4 &\rightarrow (\F{64})^2 \\
    (L_{(64)},R_{(64)},k_{0(64)},k_{1(64)})&\mapsto (FL(L,k_0),FL^{-1}(R,k_1))
\end{align*}

>>>>

\subsubsection{Encryption}

@@

\subsubsection{Key schedule}

The master key $K$ is split into two parts $K=K_{L(128)}||K_{R(128)}$ with
$K_R=0$ for 128-bit keys. Then, two variables $K_{A(128)},K_{B(128)}$ are
generated by repeated application of the round function with key constants
$\Sigma_i,i=(0,1,\dots,5)$:

\begin{align*}
    K_A&\leftarrow K_L\oplus K_R \\
    K_A&\leftarrow FE(FE(K_A, \Sigma_0=0xa09e667f3bcc908b), \Sigma_1=0xb67ae8584caa73b2) \\
    K_A&\leftarrow K_A\oplus K_L \\
    K_A&\leftarrow FE(FE(K_A, \Sigma_2=0xc6ef372fe94f82be), \Sigma_3=0x54ff53a5f1d36f1c) \\
    K_B&\leftarrow FE(FE(K_A, \Sigma_4=0x10e527fade682d1d), \Sigma_5=0xb05688c2b3e6c1fd)
\end{align*}

Subkeys are then created by rotating $K_L,K_R,K_A,K_B$:

\begin{table}[h!]
    \centering
    \scriptsize
    \caption{Subkeys for 128-bit key}
    \begin{tabular}{lll|lll}
        \toprule
        Usage & Subkey & Value & Usage & Subkey & Value \\
        \midrule
        \multirow{2}{*}{Prewhitening} & $kw_{0(64)}$ & $(K_L{\lll_0})_{L(64)}$ & $F(\text{Round 9})$ & $k_{9(64)}$ & $(K_L{\lll_{60}})_{R(64)}$ \\
                                      & $kw_{1(64)}$ & $(K_L{\lll_0})_{R(64)}$ & $F(\text{Round 10})$ & $k_{10(64)}$ & $(K_A{\lll_{60}})_{L(64)}$ \\
        $F(\text{Round 0})$ & $k_{0(64)}$ & $(K_A{\lll_0})_{L(64)}$            & $F(\text{Round 11})$ & $k_{11(64)}$ & $(K_A{\lll_{60}})_{R(64)}$ \\
        $F(\text{Round 1})$ & $k_{1(64)}$ & $(K_A{\lll_0})_{R(64)}$            & $FL$        & $kl_{2(64)}$ & $(K_L{\lll_{77}})_{L(64)}$ \\
        $F(\text{Round 2})$ & $k_{2(64)}$ & $(K_L{\lll_{15}})_{L(64)}$         & $FL^{-1}$   & $kl_{3(64)}$ & $(K_L{\lll_{77}})_{R(64)}$ \\
        $F(\text{Round 3})$ & $k_{3(64)}$ & $(K_L{\lll_{15}})_{R(64)}$         & $F(\text{Round 12})$ & $k_{12(64)}$ & $(K_L{\lll_{94}})_{L(64)}$ \\
        $F(\text{Round 4})$ & $k_{4(64)}$ & $(K_A{\lll_{15}})_{L(64)}$         & $F(\text{Round 13})$ & $k_{13(64)}$ & $(K_L{\lll_{94}})_{R(64)}$ \\
        $F(\text{Round 5})$ & $k_{5(64)}$ & $(K_A{\lll_{15}})_{R(64)}$         & $F(\text{Round 14})$ & $k_{14(64)}$ & $(K_A{\lll_{94}})_{L(64)}$ \\
        $FL$        & $kl_{0(64)}$ & $(K_A{\lll_{30}})_{L(64)}$                & $F(\text{Round 15})$ & $k_{15(64)}$ & $(K_L{\lll_{94}})_{R(64)}$ \\
        $FL^{-1}$   & $kl_{1(64)}$ & $(K_A{\lll_{30}})_{R(64)}$                & $F(\text{Round 16})$ & $k_{16(64)}$ & $(K_A{\lll_{111}})_{L(64)}$ \\
        $F(\text{Round 6})$ & $k_{6(64)}$ & $(K_L{\lll_{45}})_{L(64)}$         & $F(\text{Round 17})$ & $k_{17(64)}$ & $(K_A{\lll_{111}})_{R(64)}$ \\
        $F(\text{Round 7})$ & $k_{7(64)}$ & $(K_L{\lll_{45}})_{R(64)}$         & \multirow{2}{*}{Postwhitening} & $kw_{2(64)}$ & $(K_A{\lll_{111}})_{L(64)}$ \\
        $F(\text{Round 8})$ & $k_{8(64)}$ & $(K_A{\lll_{45}})_{L(64)}$         &                                & $kw_{3(64)}$ & $(K_A{\lll_{111}})_{R(64)}$ \\
        \bottomrule
    \end{tabular}
\end{table}

Subkey generation for 192/256-bit keys are generated in a similar way.

@@

\begin{description}
\item[Feistel round function $FE$:]

@@

\item[SP-function $F$:]

@@

\item[Substitution function $S$:]

\begin{align*}
    S:\F{64} &\rightarrow \F{64} \\
    \begin{aligned}
        l_{0(8)}||l_{1(8)}||l_{2(8)}||l_{3(8)}&||\\
        l_{4(8)}||l_{5(8)}||l_{6(8)}||l_{7(8)}&
    \end{aligned}&\mapsto
    \begin{aligned}
        s_0(l_0)||s_1(l_1)||s_2(l_2)||s_3(l_3)&||\\
        s_1(l_4)||s_2(l_5)||s_3(l_6)||s_0(l_7)&
    \end{aligned},
\end{align*}

with 8-bit S-boxes $s_0,s_1,s_2,s_3:\F{8}\rightarrow \F{8}$.

\item[Permutation function $P$:]

\begin{align*}
    P:\F{64} &\rightarrow \F{64} \\
    \begin{pmatrix}z_7\\z_6\\z_5\\z_4\\z_3\\z_2\\z_1\\z_0\end{pmatrix}&\mapsto
    \begin{pmatrix}
        0 & 1 & 1 & 1 & 1 & 0 & 0 & 1 \\
        1 & 0 & 1 & 1 & 1 & 1 & 0 & 0 \\
        1 & 1 & 0 & 1 & 0 & 1 & 1 & 0 \\
        1 & 1 & 1 & 0 & 0 & 0 & 1 & 1 \\
        0 & 1 & 1 & 1 & 1 & 1 & 1 & 0 \\
        1 & 0 & 1 & 1 & 0 & 1 & 1 & 1 \\
        1 & 1 & 0 & 1 & 1 & 0 & 1 & 1 \\
        1 & 1 & 1 & 0 & 1 & 1 & 0 & 1
    \end{pmatrix}\begin{pmatrix}z_7\\z_6\\z_5\\z_4\\z_3\\z_2\\z_1\\z_0\end{pmatrix}
\end{align*}

\item[$FL$ layer function $FLL$:]

\begin{align*}
    FLL:(\F{64})^4 &\rightarrow (\F{64})^2 \\
    (X_{L(64)},X_{R(64)},k_{0(64)},k_{1(64)})&\mapsto (FL(X_L,k_0),FL^{-1}(X_R,k_1))
\end{align*}

\item[$FL$:]

\begin{align*}
    FL:(\F{64})^2 &\rightarrow \F{64} \\
    (X_{L(32)}||X_{R(32)},k_{L(32)}||k_{R(32)})&\mapsto (Y_{L(32)}||Y_{R(32)}),
\end{align*}

where

\begin{align*}
    Y_{R(32)}&=((X_L\cap k_{L})\lll_1)\oplus X_{R} \\
    Y_{L(32)}&=(Y_R\cup k_{R})\oplus X_{L}
\end{align*}

\item[$FL^{-1}$:]

\begin{align*}
    FL^{-1}:(\F{64})^2 &\rightarrow \F{64} \\
    (Y_{L(32)}||Y_{R(32)},k_{L(32)}||k_{R(32)})&\mapsto (X_{L(32)}||X_{R(32)}),
\end{align*}

where

\begin{align*}
    X_{L(32)}&=(Y_R\cup k_{R})\oplus Y_{L} \\
    X_{R(32)}&=((X_L\cap k_{L})\lll_1)\oplus Y_{R}
\end{align*}
\end{description}
